How to Stop Useless Refactorings

### 2016-08-13 Manuscript.tex
Updates from ShareLaTeX

uu \documentclass[final,twocolumn,5p]{elsarticle}
\usepackage{tikz}
  \def\firstcircle{(90:1.75cm) circle (2.5cm)}
  \def\secondcircle{(210:1.75cm) circle (2.5cm)}
  \def\thirdcircle{(330:1.75cm) circle (2.5cm)} 
\usepackage{comment}
\usepackage{framed,graphicx}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{bigstrut}
\usepackage{color}
\usepackage{graphics} 
\usepackage{eqparbox}
\usepackage{graphics}
\usepackage{colortbl} 
%\usepackage{times}
 \usepackage{mathptmx} \usepackage[scaled=.90]{helvet} \usepackage{courier}
\usepackage{balance}
\usepackage{picture}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage[export]{adjustbox}
\renewcommand{\footnotesize}{\scriptsize}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\definecolor{lavenderpink}{rgb}{0.98, 0.68, 0.82}
\definecolor{celadon}{rgb}{0.67, 0.88, 0.69}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
%%% graph
\newcommand{\crule}[3][darkgray]{\textcolor{#1}{\rule{#2}{#3}}}

@@

\usepackage{tikz}
\usepackage{xcolor}
   \usepackage[framed]{ntheorem}
\usepackage{framed}
\usetikzlibrary{shadows}
%\newtheorem{Lesson}{Lesson}
\theoremclass{Lesson}
\theoremstyle{break}

% inner sep=10pt,
\tikzstyle{thmbox} = [rectangle, rounded corners, draw=black,
  fill=Gray!20,  drop shadow={fill=black, opacity=1}]
\newcommand\thmbox[1]{%
  \noindent\begin{tikzpicture}%
  \node [thmbox] (box){%
\begin{minipage}{.94\textwidth}%    
      \vspace{-3mm}#1\vspace{-3mm}%
    \end{minipage}%
  };%
  \end{tikzpicture}}

\let\theoremframecommand\thmbox
\newshadedtheorem{lesson}{Research answer}

@@

\title{Not Smelly Enough:  Learning What Bad Smells to Ignore}

@@

\begin{abstract} 
{\bf Context: }The SE literature lists many  reasons
for reorganizing software projects; e.g. long methods are more bug prone
so it is wise to refactoring
very large methods into numerous smaller ones.
At least some of those recommendations are outdated (based on older projects)
or irrelevant (based on other projects). Yet developers persist
is following much of that advice, without critical reviewing
its relevance to the current project.

{\bf Objective:} If a developer is doing some task for reason ``X'', 
how can that developer's manager assess if ``X'' is a good reason
that justifies that task?

{\bf Method:} For a particular task (manual refactoring of code in order
to remove a bad smell), we check   if there is any historical evidence
to endorse that refactoring.  Ysing the XTREE

bad smells that may be detrimental to a software system. But which are relevant to the current project and which can be ignored?
Developers, text books, and tools disagree on which bad smells are important. 
Further, it is unclear at what threshold values smell-related metrics
should trigger code improvement or preventative maintenance. Most studies do not offer {\em both} measurable metrics {\em and} a threshold for their proposed bad smells.

\noindent 
{\bf Objective: }We seek to propose a tool to prioritize  bad  smells.

\noindent
{\bf Method: } We introduce XTREE, a tool which, given a historical log of defects seen previously in the code, deprioritizes bad smells that have no prior association with defects.

@@

\noindent
{\bf Results: }Code modules that are changed in response to XTREE's recommendations contain significantly fewer defects than recommendations in previous studies. Further, XTREE endorses changes to far fewer code metrics, and that bad smell recommendations (learned from previous studies) are not universal to software projects.

\noindent
{\bf Conclusion: }It is important to  prioritize the importance of specific bad smells in the context of each project since not all bad smells are relevant to the current project. Based on these results, we endorse using bad smells to guiding refactoring.

@@

\noindent
``Managing people is like herding cats'', says management consultant Watler
Bennis~\cite{bennis97}. Nowhere is this more true than when managing programmers. 
Programmers can be very intelligent, but very opinionated people whose
social position within software organizations relies, in part,
of perceived notions of technical excellence. Hence, it can
be hard for managers to monitor and guide programmers away from wasting
time on nonproductive activity, particularly if that programmer can find
some support for that activity somewhere in the SE literature.

Consider, for example, the activity of refactoring code based on bad smells.
According to   Fowler ~\cite{fowler99}, bad smells (a.k.a. code smells)
are ``a surface indication that usually corresponds to a deeper problem''.
Fowler strongly recommends   removing   code smells   by
\begin{quote}
``...applying a series of small behavior-preserving transformations, each 
of which seem ``too small to be worth doing''. 
The  effect of   these transformations is quite significant. By doing them in small steps you {\em reduce the risk 
of introducing errors}'' (our italics).
\end{quote}
The advice on what  bad smells to fix is extensive and, as we show below,
contradictory (see \fig{smells}) and poorly supported. Yet
programmers often use the SE literature to justify extensive refactoring efforts.
For example, this paper began when a Washington-based software company shared
one of their management challenges with us: their  releases are delayed by developers spending  much time removing bad smells within their code. This management team wanted some way to check if all that activity was useful (hence, this article).

Note that these managers are {\em not} asking for better
automatic refactoring
tool support. Their human programmers are refactoring
code all the time, perhaps too much so.
Hence, their management goal is: 
\bi
\item
{\em Not} to optimize {\em necessary} refactorings;
\item
But rather to discourage {\em unnecessary} refactorings.
\ei
For those managers, we offer the  following heuristic:
\bi
\item If there is no evidence in the historical record that changing ``X'' is useful;
\item Then discourage developers from doing ``X''.
\ei
This paper describes one way to operationalize this heuristic, in the context
of assessing bad smells. We use data mining to generate two oracles: a {\em primary
change oracle} and a {\em secondary validation oracle}.
By combining these two oracles,
 we can generate and validate useful
changes to code modules.

@@

When these oracles were applied to defect data from
open source object-oriented software systems, we discovered a {\em disjunctive fallacy}
in the bad smell literature~\cite{erni96,bender99,Shatnawi10,Alves2010,hermans15}.

\bi
\item A common heuristic is to recommend refactoring of a code module if any static
code measure grow ``unusually  large''
(see \tion{bst} for   different ways to defining ``unusually large'').
\item This can lead to a {\em disjunctive recommendation}; i.e. given knowledge
that some static
code measures $a_i$ is ``unusually large'' after some threshold $t_i$,
then refactor code if we ever the see disjunction:
\[\left(a_1 > t_1 \right) \bigvee \left(a_2 > t_2\right) \bigvee \left(a_3 > t_3 \right)\bigvee   ...\]
\ei
This disjunctive recommendation is a fallacy.
According to XTREE's results,
there exists an intricate set of associations between static code measures
such that {\em decreasing}  $a_i$ necessitates {\em increasing} $a_j$.
It is easy to see why this is so.
A requirement for  refactoring is that the new code supports the same functionality
as before. If we pull code out of a function (since that function has grown too
large), {\em that functionality has to go somewhere else}. What we see in XTREE's change
recommendation is that if we decrease a module's lines of code  then in the refactored
code, the functionality is achieved via a communication between several different parts
of the system. That is, {\em decreasing} the lines of code metric within a module {\em increases}
that module's coupling metrics.

The XTREE results  confirmed the intuitions
of the project managers that prompted this investigation:
Yes,  indeed, their programmers
were wasting time on   performing refactorings  that  decreased multiple measures when, in fact,
they should be trying to {\em decrease} some measures while {\em increasing} others.
Also, in a study of five different JAVA OO systems,
XTREE found that  only $\frac{4}{20}=20$\% 
  of the measures were usually found within  change recommendations. That is, it is unlikely
  that refactoring based on measurements of  80\% of the code measures would be useful.

Based on the above, we recommend the use of XTREES as follows XXX

  The rest of this paper offers details on this study.  
  %YYYstructure of paper written here

  In order to support dissemination
  of research results, and to encourage other researchers to repeat our results,
  all the code used in this paper is publicly available\footntote{XXX: need a public repo}.

  \subsection{Relationship to Prior Work }
A four page prelininary report on the XTREE system~\cite{krishna2015actionable} has been presented previously\footnote{https://goo.gl/2In3Lr}. That short report offered case studies only two  of
the eight data sets studied here. Also, that prior work had
no comparison to any baseline
method from other researchers, whereas this paper compares XTREE's recommendations
with those found by researchers exploring bad smells.
Also, that paper offered
no secondary oracle, i.e. no way to check if XTREE's recommended changes were sensible.
Further, syntactically, we note that this report is four times longer than that prior paper.

XXX next text here on CD

\section{Background}
\subsection{About Bad Smells}

Bad smells are a widely used technique for evaluating code. As of  early 2016, ACM Digital Library reports 
the Fowler text of {\em Refactoring} has 6434 references~\cite{fowler99}. A search  on {\em ``bad smell''programming} 
where {\em proceedings=ICSE} finds 1,748 papers since 2006.
Hence it is important to understand the value of
this popular topic.  
Valuable development resources may be wasted on bad smells that are ``bad''; i.e. 
 irrelevant or misleading
for some particular project.

The literature is divided on the merits of bad smells.
Much research endorses them to guide
code improvement (e.g., refactoring or preventative maintenance). A recent literature review by Tufano et al.~\cite{Tufano2015}  
lists dozens of papers on smell detection and repair tools. 
Yet
other papers cast doubt on the value of bad smells
as triggers for code improvement~\cite{Mantyla2004,Yamashita2013,Sjoberg2013}. 
As we show in this paper,  
 tools, text books, and developers disagree on what bad smells
are important enough to hunt down and eliminate. Our findings
are consistent with other research suggesting that universal bad
smells that ignore project context are not useful to guide refactoring~\cite{Mantyla2004,Yamashita2013,Sjoberg2013}.

In our view, there are four major problems with bad smells.

{\bf PROBLEM \#1:} As discussed in the next section, programmer
  cognitive biases result in different
developers declaring that
bad smell X or Y is most important. 
 XTREE is useful for  assessing the relative importance of
   X vs. Y.

{\bf PROBLEM \#2:} Not all bad smells can be fixed. 
When new code must
be combined with code from a third party, 
sometimes the constraints of that code 
lead unavoidably to bad smells violations (e.g.
developers might need to passed as a single closure a  very long method
to  a third-party GUI toolkit).   In this case,
a refactoring proposed to reduce that bad smell
may be irrelevant since it cannot be applied to this code base.

Hence, XTREE {\em reasons optimistically} to find the set
$R_0$ of refactorings that \textit{might} be useful (see \tion{rq} on optimistic reasoning).
The space of \textit{actually usable} refactorings is 
the subset $R_1 \subseteq R_0$.
If a developer's proposed refactoring $R_2$ is
not in $R_0$, then it cannot be in the usable
set $R_1$.  We show experimentally that XTREE
often recommends small subsets of
code metrics as useful guides for refactoring. Thus, we
can rule out some proposed refactorings
since many of those proposals fall outside of the space
useful refactorings ($R_0$) suggested by XTREE.

{\bf PROBLEM \#3:}  
Shantnawi~\cite{Shatnawi10} warns most  studies do not offer
{\em both} a metric and a threshold for proposed bad smells. For example,  Darcy and Kemerer~\cite{darcy05}
discuss object-oriented (OO) design metrics that might defect bad smells (e.g. an excessive
depth of inheritance tree), but they do not define thresholds
for when the code starts to smell bad.  
Shatnawi~\cite{Shatnawi10}
and  Alves et al.~\cite{Alves2010} are rare cases where both metrics and thresholds are defined (see \tion{bst}).
However, their approaches do not explore the interconnections between sets of code metrics. This is important if sets of
code metrics have to be changed at the same time.
For example, it is not possible to refactor a
long method if several smaller ones without increasing
the coupling metrics of a class. 
More generally, we say that bad smell thresholds should be
{\em conjunctions} of metric+threshold pairs that need to be changed {\em together}.
Shatnawi~\cite{Shatnawi10}
and Alves et al. report {\em disjunctions} of multiple metrics+thresholds~\cite{Shatnawi10,Alves2010} without indicating how changes to these metrics are related.

To address this third problem, 
XTREE   generates {\em conjunctions} of   thresholds
than should  be considered at the same time.

{\bf PROBLEM \#4:} It is hard to judge the  effects of removing bad smells.
Refactored code cannot be assessed just by a rerun of the the test
suite (since refactorings are supposed to preserve behavior), and otherwise one must wait for new defect reports to evaluate the effect of refactoring changes on code.  
% waiting to observe 
% code refactorings is meant to
% preserve the behavior of the system (i.e. the same tests
% should pass before and after the refactoring). 
Also, it many not be practical
 to assessing refactorings by making the
refactorings,  then seeing what happens to that code for the rest
of the project. In a test-driven agile project, as
SCRUM meetings change what stories are implemented next,
the operational profile that lead to defects in the old
code may not be repeated in future executions.

To resolve this  problem, SE researchers such as 
Cheng et al.~\cite{Cheng10}, O'Keefe et al.~\cite{OKeeffe08,OKeeffe07},
Moghadam~\cite{Moghadam2011} and Mkaouer et al.~\cite{Mkaouer14}
use a pair of oracles: 
\begin{enumerate}
    \item A {\em primary oracle} system to propose changes;
\item A {\em secondary validation oracle} (which is learned separately
to the primary oracle) that   assesses
how defective is the code before and after some
refactoring.
\end{enumerate}
For the second system,
 Cheng, O'Keefe, Moghadam and  Mkaouer et al. use the QMOOD hierarchical
quality model~\cite{Bansiya02}.
A shortcoming of QMOOD
is that quality models learned from other projects
may perform poorly when applied to new projects~\cite{localvsglobal}.
Hence, for this study, we  eschew
older quality models like QMOOD. Instead, we use
Random Forests~\cite{Breiman2001} to learn defect predictors
from OO code metrics.
Note that, unlilke QMOOD, such measurements 
  are specific to the project, and the measurements can be reconstructed and the predictors rebuilt for other projects.

\subsection{Why Not Just Ask Developers to Rank Bad Smells?}\label{sect:prelim}

Why build tools like XTREE to  critique proposed developer actions?
Are not those developers the best guide to what is important or irrelevant
to their code? Developer {\em cognitive biases} can mislead them to
assert that some things are important and relevant when they are not.
This section discusses those biases-- firstly in the general sense
across all SE beliefs; then secondly in the   context of code smells.

A widespread malaise in software engineering is that
beliefs about software are rarely revised and hence may be
  inaccurate and 
misleading~\cite{passos11,jorgensen09,mei15,me16phase,prem16}. Software developers may have strong views on many issues, including bad smells, but those views may be wrong for the current
project.
Software engineering is not the only field with this problem.
  For example, the medical profession applies many practices based on studies that have been disproved (a recent article in the Mayo Clinic Proceedings~\cite{prasad13} found 146 medical practices based on studies in year i, but which were reversed by subsequent trials within years i + 10). Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}. Aschwanden warns that {\em cognitive biases} such as confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) influence how we process information~\cite{aschwanden15}.

@@

\begin{enumerate} 
% COMMENT FROM LL: This bullet is weak. Maybe goto isn't used in an unrestricted manner because of Djikstra's warning.
% \item
% Nagappan et al. analyzed~\cite{mei15} 
% how   {\tt goto}  was used  in 11,000 Github repositories. 
% They found that developers
% rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his  1968
% memo ``Go to statement considered harmful''~\cite{Dijkstra68}.
\item
In other work~\cite{me16phase} with the Software Engineering Institute (SEI), we have revisited
the ``phase delay'' truism that {\em the cost of fixing an error dramatically increases the longer it is in the system}. 
We found no evidence for such large   phase delay effect in modern
software (in 171 software projects shepherded
by the SEI, 2006-2014) possibly since it has been  heavily mitigated by $21^{st}$ century software development languages, tools and development practices. 
Whatever the reason, the main point  is that phase delay is a widely
held belief, despite little recent evidence to support it.
\item
Devanbu et al.  examined responses from 564 Microsoft software developers from around
the world, they found that  ``(a)~programmers do indeed have very
strong beliefs on certain topics; (b)~their beliefs are primarily formed
based on personal experience, rather than on findings in empirical
research; (c)~beliefs can vary with each project, but do not necessarily
correspond with actual evidence in that project''~\cite{prem16}.
\end{enumerate}
Devanbu et al. further  comment that ``programmers give personal experience
as the strongest influence in forming their opinions''. This is a troubling
result, especially given the above comments from Passos and  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, freeze, and rarely revist those opinions.

@@

@@

VARL encodes the defect count
for each class as 0 (no defects known in class) or 1 (defects known in class).
Univariate binary logistic regression is applied to learn three coefficients:  
 $\alpha$ is the intercept constant;
    $\beta$ is the coefficient for maximizing log-likelihood;
  and $p_0$  
    measures   how well this   model predicts for   defects.
Any code metric with $p_0>0.05$ is  rejected as being a poor defect predictor. Thresholds are then learned from the surviving metrics $M_c$ using
an equation proposed by Bender:
\begin{equation}
 \mathit{bad\; smell\; if\;} M_c > \frac{1}{\beta }\left( {\log \left( {\frac{{{p_1}}}{{1 - {p_1}}}} \right) - \alpha } \right) 
\end{equation}

@@

  XTREE  is a cluster delta algorithm
  that avoids the problem of verbose $\Delta$s.
  Instead of reasoning over cluster centroids,
  XTREE utilizes a decision tree learning approach
  to find the fewest differences between clusters of examples.

@@

     Once the tree is built, a class with code metric data is passed into the tree and evaluated down the tree to a leaf node (see the \textcolor{orange}{{\bf orange}} line in \fig{xtree_samp}).
      XTREE then looks for a nearby leaf node with a lower defect
      count (see the \textcolor{green}{{\bf green}} line in \fig{xtree_samp}). For that evaluated class, XTREE proposes a bad smell
      thresholds that is  the difference between 
      \textcolor{green}{{\bf green}} and \textcolor{orange}{{\bf orange}}.

@@

\input{tex/rq2.tex}
\subsection{Results}

>>>>

\documentclass[twocolumn,5p]{elsarticle}
\usepackage{tikz}
  \def\firstcircle{(90:1.75cm) circle (2.5cm)}
  \def\secondcircle{(210:1.75cm) circle (2.5cm)}
  \def\thirdcircle{(330:1.75cm) circle (2.5cm)} 
\usepackage{comment}
\usepackage{framed,graphicx}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{bigstrut}
\usepackage{color}
\usepackage{graphics} 
\usepackage{eqparbox}
\usepackage{graphics}
\usepackage{colortbl} 
%\usepackage{times}
 \usepackage{mathptmx} \usepackage[scaled=.90]{helvet} \usepackage{courier}
\usepackage{balance}
\usepackage{picture}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage[export]{adjustbox}
\renewcommand{\footnotesize}{\scriptsize}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\definecolor{lavenderpink}{rgb}{0.98, 0.68, 0.82}
\definecolor{celadon}{rgb}{0.67, 0.88, 0.69}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
%%% graph
\newcommand{\crule}[3][darkgray]{\textcolor{#1}{\rule{#2}{#3}}}

@@

\definecolor{Gray}{rgb}{0.88,1,1}
\definecolor{Gray}{gray}{0.85}
\usepackage[svgnames]{xcolor}
\usepackage[framed]{ntheorem}
\usepackage{framed}
\usepackage{tikz}
\usetikzlibrary{shadows}
%\newtheorem{Lesson}{Lesson}
\theoremclass{Lesson}
\theoremstyle{break}

% inner sep=10pt,
\tikzstyle{thmbox} = [rectangle, rounded corners, draw=black,
  fill=Gray!40,  drop shadow={fill=black, opacity=1}]
\newcommand\thmbox[1]{%
  \noindent\begin{tikzpicture}%
  \node [thmbox] (box){%
\begin{minipage}{.94\textwidth}%    
      \vspace{-2mm}#1\vspace{-3mm}%
    \end{minipage}%
  };%
  \end{tikzpicture}}

\let\theoremframecommand\thmbox
\newshadedtheorem{lesson}{Result}

@@

\title{How to Stop Useless Refactorings}

@@

\begin{abstract} 
{\bf Context: } 
Developers use code bad smells to gudie their refactoring (reorganizing their code).
Yet developers, text books, and tools disagree on which bad smells are important.
%Many bad smell detectors recommend refactoring when static code measure
%cross some threshold boundary. 
%%However, it is unclear at what threshold values smell-related metrics
%should trigger code improvement or preventative maintenance.

\noindent 
{\bf Objective: } We seek to a tool that can recognize  useless refactorings.

\noindent
{\bf Method: } We introduce XTREE, a tool which, given a historical log of defects seen previously in the code, generates
a set of useful code changes.
Any bad smell that requires changes outside of that set can be deprioritized 
(since there is no
historical evidence that that bad smell causes any problems).

@@

\noindent
{\bf Results: }Code modules that are changed in response to XTREE's recommendations contain significantly fewer defects than recommendations in previous studies. Further, XTREE endorses changes to very few code metrics, and those bad smell recommendations (learned from previous studies) are not universal to all  software projects.

\noindent
{\bf Conclusion: }
Before undertaking a refactoring based on a bad smell report,
use a tool like XTREE to check and ignore any refactoring   that  are useless; i.e. which lacks evidence
  in the historical record  that it is useful to make that change.
  Note that this use case applies to both manual refactorings proposed by developers
as well as refactoring  conducted by automatic methods.
This recommendation assumes that there is an historical record.
If none exists, then the results of this paper could be use used as a guide (see our
Table~8).

@@

\noindent
%%``Managing people is like herding cats'', says management consultant Watler
%Bennis~\cite{bennis97}. Nowhere is this more true than when managing programmers. 
%Programmers can be very intelligent, but very opinionated people whose
%social position within software organizations relies, in part,
%of perceived notions of technical excellence. Hence, it can
%be hard for managers to suggest to programmers that are
%wrong about how they are prioritizing their development effort.

%Consider, for example, the activity of refactoring code based on bad smells.
According to   Fowler ~\cite{fowler99}, bad smells (a.k.a. code smells)
are ``a surface indication that usually corresponds to a deeper problem''.
Fowler strongly recommends   removing   code smells   by
\begin{quote}
``...applying a series of small behavior-preserving transformations, each 
of which seem ``too small to be worth doing''. 
The  effect of   these \underline{refactoring} transformations is quite significant. By doing them in small steps you reduce the risk 
of introducing errors''.
\end{quote}
Bad smells are a widely used technique for evaluating code. As of  early 2016, ACM Digital Library reports 
the Fowler text of {\em Refactoring} has 6434 references~\cite{fowler99}. A search  on {\em ``bad smell''programming} 
where {\em proceedings=ICSE} finds 1,748 papers since 2006.
Hence it is important to understand the value of
this popular topic.  
Valuable development resources may be wasted on bad smells that are ``bad''; i.e. 
 irrelevant or misleading
for some particular project.

In our experience, programmers can use the SE literature to justify extensive, perhaps even excessive, refactoring efforts.
For example, this paper began when a Washington-based software company shared
one of their management challenges with us: their  releases are delayed by developers spending  much time removing bad smells within their code. This management team wanted some way to check if all that activity was useful.

Note that these managers are {\em not} asking for better
automatic refactoring
tool support. Their human programmers are refactoring
code all the time, perhaps too much so.
Hence, their management goal is: 
\bi
\item
{\em Not} to optimize {\em necessary} refactorings;
\item
But rather to stop {\em unnecessary} refactorings.
\ei
For those managers, we offer the  following heuristic:
\bi
\item If there is no evidence in the historical record that changing ``X'' is useful;
\item Then stop developers from doing ``X''.
\ei
This paper describes one way to operationalize this heuristic, in the context
of assessing bad smells. We use data mining to generate two oracles: a {\em primary
change oracle} and a {\em secondary validation oracle}.
By combining these two oracles,
 we can generate and validate useful
changes to code modules.

@@

When these oracles were applied to defect data from
open source object-oriented software systems, we discovered a {\em conjunctive fallacy}
in the bad smell literature~\cite{erni96,bender99,Shatnawi10,Alves2010,hermans15}.
A common heuristic is to recommend refactoring of ``bad'' code
to make it ``better''   as follows: for all static code measures that exceed some threshold, make changes such that the thresholds are no longer exceeded. That is:
\begin{equation}\label{eq:df}
\scriptstyle
\begin{array}{rl}
\mathit{bad} & = \left(a_1 > t_1 \right) \bigvee \left(a_2 > t_2\right) \bigvee    ... \\
\mathit{better} & = \neg\;\mathit{bad} = \left(a_1 \le t_1 \right) \bigwedge \left(a_2 \le t_2\right)  \bigwedge  ...
\end{array}
\end{equation}
We say that the above definition of ``better'' is a conjunctive fallacy
since it assumes that the best way to improve code is to decreases multiple code attribute measures below $t_i$ in order
to remove the ``bad'' smells. This is wrong since 
there exists an intricate set of associations between static code measures
such that {\em decreasing}  $a_i$ necessitates {\em increasing} $a_j$.
It is easy to see why this is so.
A requirement for  refactoring is that the new code supports the same functionality
as before. If we pull code out of a function (since that function has grown too
large), {\em that functionality has to go somewhere else}. What we see in XTREE's change
recommendation is that if we decrease a module's lines of code  then in the refactored
code, the functionality is achieved via a communication between several different parts
of the system. That is, {\em decreasing} the lines of code metric within a module {\em increases}
that module's coupling metrics.

  The rest of this paper is structured as follows. First, we offer some background on bad smells. After that,
  in Section~3, 
  different methods for determining when code does/does not have bad smell. One of those methods
  will be XTREE while the others will come from other researchers. Next,    Section~4
  defines an assessment method that checks which of these methods is most useful. The results from that assessment 
  are presented in Section~5. Those results will answer five research questions.

 {\bf  RQ1: Effectiveness}: According to the validation oracle, which of the methods  defined in Section~3 is the best  change oracle for identifying what and how
  code modules should be changed? To answer this question, we used data from five OO Java projects
  (Ivy, Lucene, Ant, Poi, Jedit). It was found that:
  \begin{lesson}
 XTREE is the best oracle on how to change code modules (in order to reduce defects).
  \end{lesson}

 {\bf RQ2: Succinctness}: Our goal is to critique and, possibly,
ignore irrelevant bad smell detectors.  If those changes are minimal (i.e. effect fewest attributes) then those changes
will be easiest to apply and monitor. Hence, we ask, which of the Section~3 methods recommended changes to the fewest
code attributes?
\begin{lesson}
Of all the code change oracles studied here, XTREE recommend  far fewest changes to static code measures.
\end{lesson}

 {\bf RQ3: Stopping}: Our advice to managers is to stop refactoring based on changes that lack
historical evidence of being effective. How effective is XTREE at offering such ``stopping points'' (i.e. clear guidance on what {\em not} to do)??
\begin{lesson}
In  any  project,  XTREE’s  recommended  changes  only  mentions one to four 
of the  static code attributes.  Any bad smell defined in terms of the remaining 19 to 16 code attributes (i.e. most of them)
would hence be deprecated.
\end{lesson}

 {\bf RQ4: Stability}: Across different projects, how variable are the changes recommended by our best change oracle?
\begin{lesson}
The direction of change recommended  XTREE is  very stable across repeated runs of the program 
\end{lesson}

{\bf RQ5: Conjunctive Fallacy}:  
Is it always  useful  to aplly \eq{df}; i.e. make code better by   reducing the values of  multiple code attributes? We find that:
\begin{lesson}
XTREE usually recommends reducing lines of code (size of the modules).
That said,  XTREE often recommends {\em increasing} the values of other static code attributes.
\end{lesson} 
Note that {\bf RQ3, RQ4, RQ5} 
confirms the intuitions
of the project managers that prompted this investigation:
\bi
\item
Yes,  indeed, their programmers
were wasting time on   performing refactorings  that  decreased multiple measures when, in fact,
they should be trying to {\em decrease} some measures while {\em increasing} others.
\item
Across all studied projects, XTREE found that  around one-fifth  
  of the measures were usually found within  change recommendations. That is, it is 80\% likely that
  that refactoring based on reducing any one measurement picked at random will be useful.
\ei
Consequently, we  recommend that XTREE be used as follows:
\bi
\item Before doing  refactoring based on a bad smell report...
\item ...check and stop any refactoring   for which there is no proof
  in the historical record that that change improves the code.
\ei
Note that this use case applies to both manual refactorings proposed by developers
as well as refactoring as conducted by automatic methods~\cite{mkaouer2015many}.
Note also that
this recommendation assumes that there is an historical record. If none exists, then the results of this paper could be use used as a guide (see our Table 8).

 \subsection{Replication Package}

  In order to support dissemination
  of research results, and to encourage other researchers to repeat our results,
  all the code used in this paper is publicly available\footnote{XXX: need a public repo}.

  \subsection{Relationship to Prior Work }
A four page prelininary report on the XTREE system~\cite{krishna2015actionable} has been presented previously\footnote{https://goo.gl/2In3Lr}. That short report offered case studies only two  of
the five data sets studied here. Also, that prior work had:
\bi
\item
no comparison to any baseline
method from other researchers, whereas this paper compares XTREE's recommendations
with those found by researchers exploring bad smells.
\item
no {\em secondary validation oracle}, i.e. no way to check if XTREE's recommended changes were sensible.
\ei
Further, of this entire paper, the only sections containing material
found in prior papers is 3.2.1, 3.2.2 as well as two-fifths of \fig{}.
%YYY section1,2, 3.1,4,
% Prior to XTREE, the authors worked on
% the CD algorithm discussed in ~\tion{cdcd}. 
% The goal of CD was the same as WHERE-- find recommendations on how to best change software
% modules. However, as discussed in~\tion{cdcd},
% CD worked very different to WHERE. Further, as shown in this paper,
% CD is significantly out-performed by XTREE.

\section{Why Not Just Ask Developers to Rank Bad Smells?}\label{sect:prelim}

Why build tools like XTREE to  critique proposed developer actions?
Our answer, documented in this section, is that (1)~SE literature is not clear which bad smells are ignorable or  high priority and must be resolved.
Also, (2)~developers themselves are unclear on what bad smells are most important to fix.

Much research endorses them to guide
code improvement (e.g., refactoring or preventative maintenance). A recent literature review by Tufano et al.~\cite{Tufano2015}  
lists dozens of papers on smell detection and repair tools. 
Yet
other papers cast doubt on the value of bad smells
as triggers for code improvement~\cite{Mantyla2004,Yamashita2013,Sjoberg2013}. 
As we show in this paper,  
 tools, text books, and developers disagree on what bad smells
are important enough to hunt down and eliminate. Our findings
are consistent with other research suggesting that universal bad
smells that ignore project context are not useful to guide refactoring~\cite{Mantyla2004,Yamashita2013,Sjoberg2013}.

If the SE literature is contradictory, why not ignore it and use domain experts (software engineers) to decide
what bad smells to fix? We do not recommend this since developer {\em cognitive biases} can mislead them to
assert that some things are important and relevant when they are not. 
A widespread malaise in software engineering is that
beliefs about software are rarely revised and hence may be
  inaccurate and 
misleading~\cite{passos11,jorgensen09,mei15,me16phase,prem16}. Software developers may have strong views on many issues, including bad smells, but those views may be wrong for the current
project.
Software engineering is not the only field with this problem.
  For example, the medical profession applies many practices based on studies that have been disproved (a recent article in the Mayo Clinic Proceedings~\cite{prasad13} found 146 medical practices based on studies in year i, but which were reversed by subsequent trials within years i + 10). Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}. Aschwanden warns that {\em cognitive biases} such as confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) influence how we process information~\cite{aschwanden15}.

@@

\bi
% COMMENT FROM LL: This bullet is weak. Maybe goto isn't used in an unrestricted manner because of Djikstra's warning.
% \item
% Nagappan et al. analyzed~\cite{mei15} 
% how   {\tt goto}  was used  in 11,000 Github repositories. 
% They found that developers
% rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his  1968
% memo ``Go to statement considered harmful''~\cite{Dijkstra68}.
\item
In other work~\cite{me16phase} with the Software Engineering Institute (SEI), we have revisited
the ``phase delay'' truism that {\em the cost of fixing an error dramatically increases the longer it is in the system}. 
We found no evidence for such large   phase delay effect in modern
software (in 171 software projects shepherded
by the SEI, 2006-2014) possibly since it has been  heavily mitigated by $21^{st}$ century software development languages, tools and development practices. 
Whatever the reason, the main point  is that phase delay is a widely
held belief, despite little recent evidence to support it.
\item
Devanbu et al.  examined responses from 564 Microsoft software developers from around
the world, they found that  ``(a)~programmers do indeed have very
strong beliefs on certain topics; (b)~their beliefs are primarily formed
based on personal experience, rather than on findings in empirical
research; (c)~beliefs can vary with each project, but do not necessarily
correspond with actual evidence in that project''~\cite{prem16}.
\ei
Devanbu et al. further  comment that ``programmers give personal experience
as the strongest influence in forming their opinions''. This is a troubling
result, especially given the above comments from Passos and  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, freeze, and rarely revist those opinions.

@@

Having made the case for automatic support for assessing bad smells,
this section reviews different ways for building those tools (one of those tools,
XTREE, will be our recommended {\em primary change oracle}).
Later in this paper, we will offer  a {\em secondary validation oracle}
that checks the effectiveness of the changes proposed by XTREE.

@@

VARL encodes the defect count
for each class as 0 (no defects known in class) or 1 (defects known in class).
Univariate binary logistic regression is applied to learn three coefficients:  
 $\alpha$ is the intercept constant;
    $\beta$ is the coefficient for maximizing log-likelihood;
  and $p_0$  
    measures   how well this   model predicts for   defects.
Any code metric with $p_0>0.05$ is  ignored as being a poor defect predictor. Thresholds are then learned from the surviving metrics $M_c$ using
an equation proposed by Bender:
\begin{equation}
 \mathit{bad\; smell\; if\;} M_c > \frac{1}{\beta }\left( {\log \left( {\frac{{{p_1}}}{{1 - {p_1}}}} \right) - \alpha } \right) 
\end{equation}

@@

  XTREE  is a cluster delta algorithm
  that avoids the problem of verbose $\Delta$s.
  It is our recommended {\em primary change oracle} that makes recommendations 
  of what changes should be made to code modules.
  Instead of reasoning over cluster centroids,
  XTREE utilizes a decision tree learning approach
  to find the fewest differences between clusters of examples.

@@

     Once the tree is built, a class with code metric data is passed into the tree and evaluated down the tree to a leaf node (see the \textcolor{orange}{{\bf orange}} line in \fig{xtree_samp}).
      XTREE then looks for a nearby leaf node with a lower defect
      count (see the \textcolor{green}{{green}} line in \fig{xtree_samp}). For that evaluated class, XTREE proposes a bad smell
      thresholds that is  the difference between 
      \textcolor{green}{{green}} and \textcolor{orange}{{\bf orange}}.

@@

\input{tex/rq2.tex}
\section{Results}
